---
title: "Advanced Usage of livCH4ipcc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage of livCH4ipcc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(livCH4ipcc)
```

## Advanced Usage of livCH4ipcc

This vignette covers advanced usage scenarios and customization options for the livCH4ipcc package.

### Customizing Input Parameters

The `User_Input` class accepts a wide range of parameters to customize your emission calculations:

```{r eval=FALSE}
# Create a User_input object with detailed parameters
input <- User_Input$new(
  animal_type = "dairy_cattle",
  animal_number = 100,
  weight = 650,
  age = 4,                    # Age in years
  gender = "female",          # Gender of animals
  reproduction = TRUE,        # Whether animals are used for reproduction
  manure_excretion = 21900,   # kg/(head.yr)
  manure_cn = 15,             # C:N ratio of manure
  fraction_housing = 0.8,     # Fraction of time in housing
  fraction_grazing = 0.2,     # Fraction of time grazing
  fraction_actual_grass = 0.15, # Actual days on grass considering climate
  feed_intake = 7300,         # kg DM/(head.yr)
  fraction_diet_grass = 0.3,  # Fraction of grass in diet
  fraction_diet_beet = 0.05,  # Fraction of beet in diet
  max_manure_CH4 = 0.24       # m3 CH4/kg VS
)
```

### Debug Mode

You can enable debug mode to get more verbose output during calculations:

```{r eval=FALSE}
# Create Inventory object with debug mode enabled
inv <- Inventory$new(input, debug_mode = TRUE)
results <- inv$run_inventory()
```

### Accessing Detailed Results

The results object contains detailed information about emissions from different sources:

```{r eval=FALSE}
# Access enteric fermentation emissions
enteric_emissions <- results$emissions$ch4_enteric

# Access manure management emissions by source
housing_emissions <- results$emissions$ch4_manure$housing
grazing_emissions <- results$emissions$ch4_manure$grass

# Access total emissions
total_emissions <- results$emissions$ch4_total

# Access emission factors used in calculations
emission_factors <- results$emission_factors
```

### Customizing Configuration Files

The package uses YAML configuration files to store default emission factors and parameters. You can customize these files or create your own:

1. First, locate the default configuration files:

```{r eval=FALSE}
system.file("extdata/enteric_fermentation.yaml", package = "livCH4ipcc")
system.file("extdata/global_parameters.yaml", package = "livCH4ipcc")
system.file("extdata/manure_management.yaml", package = "livCH4ipcc")
```

2. Copy and modify these files to create your custom configurations

3. Use your custom configuration files when creating an Inventory object:

```{r eval=FALSE}
# Create an Inventory object with custom configuration paths
inv <- Inventory$new(input)
inv$config_paths <- list(
  enteric_fermentation = "path/to/custom/enteric_fermentation.yaml",
  global_parameters = "path/to/custom/global_parameters.yaml",
  manure_management = "path/to/custom/manure_management.yaml"
)

# Run inventory with custom configurations
results <- inv$run_inventory()
```

### Working with Multiple Animal Types

You can calculate emissions for multiple animal types and combine the results:

```{r eval=FALSE}
# Create inputs for different animal types
dairy_input <- User_Input$new(animal_type = "dairy_cattle", animal_number = 100, ...)
beef_input <- User_Input$new(animal_type = "beef_cattle", animal_number = 200, ...)
pig_input <- User_Input$new(animal_type = "pigs", animal_number = 500, ...)

# Calculate emissions for each animal type
dairy_results <- Inventory$new(dairy_input)$run_inventory()
beef_results <- Inventory$new(beef_input)$run_inventory()
pig_results <- Inventory$new(pig_input)$run_inventory()

# Combine total emissions
total_farm_emissions <- dairy_results$emissions$ch4_total + 
                        beef_results$emissions$ch4_total + 
                        pig_results$emissions$ch4_total
```

### Seasonal Calculations

For dairy cattle, you can access seasonal emission factors:

```{r eval=FALSE}
# Create an Inventory object
inv <- Inventory$new(input)

# Calculate seasonal emission factors
seasonal_ef <- inv$calculate_seasonal_ef()

# Access winter and summer emission factors
winter_ef <- seasonal_ef$winter_ef
summer_ef <- seasonal_ef$summer_ef
```

## Extending the Package

The package is built using R6 classes, which makes it easy to extend with your own custom functionality:

```{r eval=FALSE}
# Create a custom Inventory class that extends the base class
CustomInventory <- R6::R6Class("CustomInventory",
  inherit = Inventory,
  public = list(
    # Add custom methods
    calculate_carbon_footprint = function() {
      # Convert CH4 emissions to CO2 equivalent (GWP of CH4 = 28)
      co2e <- self$results$emissions$ch4_total * 28
      return(co2e)
    }
  )
)

# Use the custom class
custom_inv <- CustomInventory$new(input)
results <- custom_inv$run_inventory()
carbon_footprint <- custom_inv$calculate_carbon_footprint()
```
