---
title: "Introduction to livCH4ipcc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to livCH4ipcc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(livCH4ipcc)
```

## Introduction

livCH4ipcc is an R package for calculating methane (CH4) emissions from livestock based on IPCC methodologies. It supports both Tier 1 and Tier 2 approaches for enteric fermentation and manure management, with seasonal variations and animal-specific parameters.

## Basic Usage

The package is built around two main classes:

1. `User_Input`: For defining animal characteristics and management practices
2. `Inventory`: For performing the emission calculations

Here's a basic example of how to use the package:

```{r eval=FALSE}
# Create a User_input object for dairy cattle
input <- User_Input$new(
  animal_type = "dairy_cattle",
  animal_number = 100,
  weight = 650,
  manure_excretion = 21900,
  manure_cn = 15,
  fraction_housing = 0.8,
  fraction_grazing = 0.2,
  feed_intake = 7300,
  fraction_diet_grass = 0.3,
  fraction_diet_beet = 0.05
)

# Create an Inventory object and run calculations
inv <- Inventory$new(input)
results <- inv$run_inventory()

# Access the results
print(results$emissions$ch4_total)
```

## Available Animal Types

The package supports various livestock types, including:

- dairy_cattle
- beef_cattle
- pigs
- sheep
- goats
- poultry

## Tier Levels

The package implements two methodological tiers according to IPCC guidelines:

- **Tier 1**: Uses default emission factors based on animal type
- **Tier 2**: Uses more detailed animal-specific parameters to calculate emissions

The appropriate tier is automatically selected based on the animal type and available input parameters.

## Seasonal Variations

For certain animal types (e.g., dairy cattle), the package can calculate seasonal emission factors for winter and summer conditions, based on differences in diet composition and management practices.

## Configuration Files

The package uses YAML configuration files to store default emission factors and parameters. These are located in the `inst/extdata` directory and can be customized if needed.

## Next Steps

For more advanced usage and customization options, see the "Advanced Usage" vignette.
